### 2021-06-29 abs.tex
got some stuff figured out .

>>>>

Bedload transport is the movement of coarse sediment grains in a river channel by rolling, sliding, and bouncing.
Motivated by the need to predict river morphodynamics, a central problem in river science is to calculate the downstream flux of grains moving as bedload.
This problem is usually formulated in terms of continuum mechanics, but this approach is questionable considering that bedload grains are coarse, they rarely reach densities approximating a continuum.
This thesis contains four projects that abandon the continuum hypothesis to interpret sediment transport in terms of individual grains.
First, I develop a new theory to calculate the sediment flux from the trajectories of individual grains as they alternate between motion and rest.
This leads to a sediment flux probability distribution which predicts mean sediment fluxes that depend on the timescale over which they are observed, explaining why measurements of the bedload flux can give such poor results.
Second, I evaluate the interplay between bed elevation changes and sediment transport rates, finding that because aggradation and degradation oppose changes in the sediment flux, bed elevation changes buffer sediment transport fluctuations.
Third, I incorporate the process of sediment burial into a model of downstream sediment trajectories, demonstrating that burial eventually halts downstream transport.
This has implications for the transport of contaminants in river channels.
Finally, using concepts borrowed from the theory of granular gases, I develop a stochastic model of individual particles moving downstream in the turbulent flow through a sequence of collisions with the bed surface. This indicates that the velocity characteristics of sediment particles are tuned by the amount of dissipation particles experience when colliding with the sedimentary bed.
This research has implications across river science, where there is increasing acknowledgement of the shortcomings of the continuum hypothesis for predicting dynamics in mountain streams. This work extends a wide body of work on individual particle motions which originates from the 1930s, and unifies it with more recent approaches to calculate the bulk bedload flux which have not, until now, been related to the movement characteristics of individual grains.

\endinput
